---
title    : ETF Correlations
subtitle : Sniny app to display ETF correlations
author   : Bill Kelemen
date     : "February 5, 2017"
output   : 
  slidy_presentation:
    highlight: pygments
    font_adjustment: -1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# source shiny app global variables and utility functions
source('../Shiny-app/stock_corr.R', echo=FALSE)
```

## Summary

This Shiny App allows a user to analyze the correlation between a sector ETF and an Index or another sector ETF. The app can be run from [here](https://bkelemen.shinyapps.io/StockCorrelationShinyApp/).

An [ETF](https://en.wikipedia.org/wiki/Exchange-traded_fund) is an investment fund traded on stock exchanges, much like stocks. The rolling 20-day correlation with an index or another ETF allows us to understand how the ETF returns are correlated to the index.

Example, for ETF sector "Consumer Discretionary" and Index "SPY Index", the chart below indicates that on average the correlation is close to 0.9 (highly correlated):
```{r genchart, echo=FALSE, cache=TRUE, fig.width=8, fig.height=3, fig.align='center'}
load_data() %>%
  isolate_index(get_ticker("SPY Index")) %>%
  calc_correlations() %>%
  get_chart("SPY Index", "Consumer Discretionary")
```

## Why is correlation important?

When holding a portfolio of stocks, ETFs and other securities, we should take into account the correlation between them. This is important for the concept of *diversification*[1], that when properly applied, helps to limit our risk.

Having a portfolio of only one stock seems very risky, as we're essentially betting all our investment on one stock. A better way is to have more than one stock ([remember  Enron?](https://en.wikipedia.org/wiki/Enron_scandal)). Doing so then raises the question of which stocks?

If we choose stocks highly correlated, then our portfolio will move in tandem to them, and we have not largely reduced our risk.

If we choose stocks that are not correlated or inversely correlated, then a decline in one will not necessarily mean a decline on the others. Therefore when our portfolio is well diversified, we should on average get a better return at lower risk than any individual security in the portfolio.

## How do we calculate correlation?

These calculations will rely on the `R` package `tidyquant`.

To calculate the correlations between ETFs and the selected Index we follow this process which is a modified version of ["Reproducible Finance With R: Sector Correlations"](https://www.r-bloggers.com/recreating-rviews-reproducible-finance-with-r-sector-correlations/)[2].

1. Download ETF prices for our sectors and calculate the [log of the returns](https://quantivity.wordpress.com/2011/02/21/why-log-returns/)[3]
2. Isolate the Index returns to which all other ETFs will be correlated
3. Calculate rolling 20-day correlations

The source code is available on this [github repo](https://github.com/bkelemen56/StockCorrelationShinyApp).

## References

[1] Diversification: Protecting Portfolios From Mass Destruction, http://www.investopedia.com/articles/financial-theory/09/uncorrelated-assets-diversification.asp

[2] Recreating RView’s "Reproducible Finance With R: Sector Correlations", mattdancho.com, https://www.r-bloggers.com/recreating-rviews-reproducible-finance-with-r-sector-correlations/

[3] Why Log Returns, https://quantivity.wordpress.com/2011/02/21/why-log-returns/

